### 2020-10-05 2020/WCGO_MCO_OptimizationLetters/2020_WCGO_MCO.tex
intermidiate version of answers

%\usepackage{hyperref}
%\usepackage{academicons}
%\RequirePackage[colorlinks,citecolor=blue,urlcolor=blue]{hyperref}
%
% \usepackage{mathptmx}      % use Times fonts if available on your TeX system
%
% insert here the call for the packages your document requires
%\usepackage{latexsym}
% etc.
%
% please place your own definitions here and don't use \def but
% \newcommand{}{}
%
% Insert the name of "your journal" with
\journalname{Optimization Letters}
%
\begin{document}

@@

The further structure of the paper is as follows. Section \ref{sec:2} presents a new class of optimization problems–multistage lexicographic multicriteria optimization (MMCOlex) problems, the solution of which is reduced to solving a sequence of global optimization problems with nonlinear constraints. Section \ref{sec:3} considers the basics of the developed approach: reducing multidimensional MCOlex problems to one-dimensional global optimization problems, applying efficient global search algorithms developed in the framework of the information-statistical theory of multiextremal optimization, and reusing the search information obtained in the process of calculations. Section \ref{sec:4} provides a theoretical analysis of the effectiveness of reusing the search information for solving MMCOlex problems. Section \ref{sec:5} contains the results of numerical experiments. Discussion of the results and suggestions for further research conclude the paper.

@@

It should also be noted that during the calculation process, it may be necessary to change the selected values of the feasible concessions $\delta_i$, $1 \leq i \leq s$, the concessions may be quite rigid or, conversely, excessively large. In general, the change in the order of efficiency criteria importance may be required. Such assumptions necessitate a more general formulation of how to solve the MCOlex problem and to provide possible solutions to many problems of the form (\ref{eq:9})
\begin{equation}\label{eq:10}
\mathbb{P}_t=\{ P_1,P_2,\dots,P_t\},
\end{equation}
which can be changed dynamically during calculations by adding new or deleting existing multicriteria optimization problems. In general, this approach defines a new class of optimization problems: multistage lexicographic multicriteria optimization problems (MMCOlex).

@@

As noted earlier, the MCOlex problem at each stage (\ref{eq:9}) of the $\varepsilon$-constraint method is a global optimization problem with nonlinear constraints (\ref{eq:16}). Generally, the solution to such problems should start from the very beginning; and multiple solutions to such global optimization problems may require extensive calculations. However, the presence of search information $A_k$ from (\ref{eq:21}) allows bringing the results of previous calculations to the values of any next solved optimization problem in the scheme (\ref{eq:9}) without any time-consuming calculations of the values of the criteria $f_i (y)$, $1 \leq j \leq s$, of the problem (\ref{eq:1}) \cite{c30,c31}. This reuse of search information to find the next efficient decision reduces the number of computations needed to solve each next optimization problem to just a few iterations -- see Section \ref{sec:5} for the results of numerical experiments.

Moreover, using the search information $A_k$ allows reducing the multistage solution of the MCO problem in accordance with the scheme (\ref{eq:9}) to the solution of the only global optimization problem in the last stage of the scheme (\ref{eq:9}), namely
\begin{equation}\label{eq:30}
P_{\delta lex} (f,\delta,D)=Arg \min_{y \in D} {f_s(y)}, f_j(y)\leq f_j^* + \delta_j, 1 \leq j < s,
\end{equation}
where for the unknown a priori values $f_j^*$, $1 \leq j < s$, the estimates $z_j^k$, ${1 \leq j < s}$, from (\ref{eq:23}) obtained from the search information $A_k$ can be used, and for the concessions $\delta_j$, $1 \leq j < s$, the normalized values can be applied, that is
\begin{equation}\label{eq:31}
	\begin{matrix}
	\delta_j = z_j^k + \theta_j (g_j^k-z_j^k ), 0 \leq \theta_j \leq 1, 1 \leq j < s,\\
	g_j^k = \max{\{ g_j (x_i ) : \nu(x_i)=j, 1 \leq i \leq k \}}.
	\end{matrix}
\end{equation}

In particular, a similar one-step scheme (\ref{eq:30})--(\ref{eq:31}) can be applied through the use of AGCS.

With a complete scheme for calculating criteria values, i.e. when at the iteration points $x_i$, $1 \leq i \leq k$, the values of all criteria $f_j(x_i)$, $1 \leq j \leq s$, $1 \leq i \leq k$, are calculated, the effect of using the search information $A_k$ can be even more significant. In this case, the search information can also be used when changing the values of the concessions $\delta_j$, $1 \leq j < s$, and thus, the solution of the next problem from the $\mathbb{P}_t$ family (\ref{eq:10}) can be carried out repeatedly using the results of all previously performed calculations. The results of numerical experiments (see Section \ref{sec:5}), demonstrate that the volume of computations performed, based on reusing search information, can significantly decrease.

The AGCS algorithm, supplemented by the ability to use search information to solve multiple MCOlex problems, will be referred to below as the global search algorithm for multistage solutions of a set of multiextremal optimization problems with nonlinear constraints (MAGCS).

@@

For MAGCS, 50 subproblems (\ref{eq:30}) were solved for different values $\theta=\theta_1$ from (\ref{eq:31}), uniformly distributed in the interval $[0,1]$. The accuracy $\varepsilon=0.05$ from (\ref{eq:28}) and the reliability parameters $r_1=r_2=3.0$ from (\ref{eq:29}) were used. The results of the executed experiments are presented in Table \ref{tab:1}.

@@

All the subsequent numerical experiments focused on the solution of 100 two-criteria MCO problems, the efficiency criteria of which were set by the GKLS generator. For each MCO problem from the generated set, from 10 to 100 subproblems (\ref{eq:30}) were solved for different values $\theta=\theta_1$ from (\ref{eq:31}), uniformly distributed in the interval $[0,1]$. The obtained results of numerical experiments were averaged by the number of solved MCO problems.

The solution of MCO problems was carried out before the stopping criterion of (\ref{eq:28}) was fulfilled, and the MCO problems were considered solved correctly if the distance of the decisons to the Pareto set did not exceed the given accuracy $\varepsilon$ from (\ref{eq:28}).

@@

The first column (SP) in the Table \ref{tab:2} contains the total number of MCO subproblems to be solved. The PI columns show the average number of iterations for solving one MCO problem. The SPI columns represent the average number of iterations for solve a single MCO subproblem. The SPS columns contain a fraction of MCO subproblems solved with the required accuracy. The HV and DU columns show the hypervolume (HV) and the distribution uniformity (DU) indices of the calculated approximation of the Pareto set, respectively. The last column (RI) shows the degree of the reduction in the number of global search iterations performed when solving subproblems (\ref{eq:30}) by reusing the search information.

The results presented in Table \ref{tab:2} demonstrate that the use of search information allows to obtain an estimate of the Pareto set after a substantially smaller total number of iterations of the global search (the PI columns). It should also be noted that as the number of solved subproblems increases, the MCO allows to build estimates of the Pareto set with better HV and DU indices.

For a better visual representation, Fig. \ref{fig:4} shows the average number of iterations performed to solve a single MCO subproblem with and without search information. The presented graphs convincingly demonstrate a significant reduction in the number of global search iterations when solving the MCO subproblems using search information.

\begin{figure}
  \centering
  \includegraphics[width=1\linewidth]{fig4}
  \caption{Average number of iterations required to solve one MCO subproblem depending on the total number of subproblems solved}
  \label{fig:4}
\end{figure}

Fig. \ref{fig:5} shows the reduction in the number of global search iterations when solving MCO subproblems using search information. The graphs presented show that when the number of solved MCO subproblems grows linearly, the number of iterations required to solve one subproblem is reduced almost linearly as well.

\begin{figure}
  \centering
  \includegraphics[width=1\linewidth]{fig5}
  \caption{Reduction of the number of iterations performed to solve one MCO subproblem depending on the total number of subproblems solved}
  \label{fig:5}
\end{figure}

It should be noted that the number of iterations of the global search is also reduced in the course of sequential solution of subproblems of the same MCO problem.

It should be noted that the number of global search iterations is also reduced in the course of sequential solution of the subproblems of the same MCO problem. Table \ref{tab:3} illustrates this trend showing the average number of iterations required to find a solution to one MCO subproblem, with subdividing the sequence of solved subproblems in groups of 10 MCO subproblems. The first row (SP) contains the MCO subproblem numbers belonging to the individual subproblem groups. The PI rows show the total number of iterations of the global search performed for all previous problems. The rows labeled as ``PI new'' show the number of iterations performed to solve the next group of MCO subproblems with the numbers of the problems given in each column heading. The last row labeled as ``FPI new'' shows the proportion of new iterations of global search performed to solve the next group of MCO subproblems.

@@

Fig. \ref{fig:6} shows the values of the achieved reduction in the number of additional iterations required to solve the next set of subproblems (the last row of Table \ref{tab:3}). The above data demonstrates that solving the next group of MCO subproblems requires about $6\%$ of new global search iterations in addition to the available search information.

\begin{figure}
  \centering
  \includegraphics[width=1\linewidth]{fig6}
  \caption{Proportion of new global search iterations for solving the next group of MCO subproblems}
  \label{fig:6}
\end{figure}

The third set of numerical experiments was designed to demonstrate the scalability of the proposed approach by solving more complex two-criteria four-dimensional MCO problems under the same conditions as in the previous series of calculations. The results of the experiments are presented in Tables \ref{tab:4}--\ref{tab:5} and in Fig. \ref{fig:4}--\ref{fig:6}. These data demonstrate that the experiments carried out to solve more complex MCO problems confirm the conclusions made earlier -- for example, solving the next group of MCO subproblems requires, in addition to the available search information, from $0.5\%$ to $14.2\%$ more global search iterations (Table \ref{tab:5}, last row).

@@

In this paper a new approach is proposed for solving computationally expensive lexicographic multicriteria optimization problems (MCOlex), which can be complicated by the multiextremal nature of the efficiency criteria and extensive volume of computations required to calculate the criteria values. A key feature of this class of problems is the ability to alter the ranking of efficiency criteria in terms of importance during calculations, which in turn necessitates a multistage solution for MCOlex problems.

The computational complexity of addressing the formulated new class of MCOlex problems can be overcome by solving a sequence of global optimization problems with nonlinear constraints using efficient information-statistical methods of global optimization. A core element in this approach is the ability to use all the search information obtained in the course of solving the multistage MCOlex problems. With each new stage of the solution, this search information allows incorporating the previously calculated values of efficiency criteria into the values of the next scalar problem of multiextremal optimization. The search information retrieved becomes part of the current optimization state and, through optimization methods, results in adaptive planning of the iterations to run the global search.

The results of numerical experiments demonstrate that this approach can sufficiently reduce the computational complexity of multistage MCOlex problem-solving.

>>>>

\usepackage{hyperref}
\usepackage{cite}

%\usepackage{academicons}
%\RequirePackage[colorlinks,citecolor=blue,urlcolor=blue]{hyperref}
%
% \usepackage{mathptmx}      % use Times fonts if available on your TeX system
%
% insert here the call for the packages your document requires
%\usepackage{latexsym}
% etc.
%
% please place your own definitions here and don't use \def but
% \newcommand{}{}
%
% Insert the name of "your journal" with
\journalname{Optimization Letters}
%
\begin{document}

@@

The further structure of the paper is as follows. Section \ref{sec:2} presents a new class of optimization problems -- multistage lexicographic multicriteria optimization (MMCOlex) problems, the solution of which is reduced to solving a sequence of global optimization problems with nonlinear constraints. Section \ref{sec:3} considers the basics of the developed approach: reducing multidimensional MCOlex problems to one-dimensional global optimization problems, applying efficient global search algorithms developed in the framework of the information-statistical theory of multiextremal optimization, and reusing the search information obtained in the process of calculations. Section \ref{sec:4} provides a theoretical analysis of the effectiveness of reusing the search information for solving MMCOlex problems. Section \ref{sec:5} contains the results of numerical experiments. Discussion of the results and suggestions for further research conclude the paper.

@@

It should also be noted that during the calculation process, it may be necessary to change the selected values of the feasible concessions $\delta_i$, $1 \leq i \leq s$, the concessions may be quite rigid or, conversely, excessively large. In general, the change in the order of efficiency criteria importance may be required. Such assumptions necessitate a more general formulation of how to solve the MCOlex problem and to provide possible solutions to many problems of the form (\ref{eq:9})
\begin{equation}\label{eq:10}
%\mathbb{P}_t=\{ P_1,P_2,\dots,P_t\},
\mathbb{F}=\{ F_1,F_2,\dots,F_t\},
\end{equation}
%new text begin
where $F_t, t>1,$ is a MCOlex problem. The size of the set $\mathbb{F}$ can be changed 
dynamically by adding new required MCOlex problems or removing unnecessary ones. 
The problems of the set $\mathbb{F}$ specify various requirements for the desired solutions and, thus, all problems of the set $\mathbb{F}$ should be solved in the course of calculations.
In general, this approach defines a new class of optimization problems: multistage lexicographic multicriteria optimization problems (MMCOlex)\footnote{From a general point of view, the MCOlex problem can be considered as a special case of the MMCOlex problem for $t=1$.}.
%new text end

@@

As noted earlier, the MCOlex problem at each stage (\ref{eq:9}) of the $\varepsilon$-constraint method is a global optimization problem with nonlinear constraints (\ref{eq:16}). Generally, the solution to such problems should start from the very beginning; and multiple solutions to such global optimization problems may require extensive calculations. However, the presence of search information $A_k$ from (\ref{eq:21}) allows bringing the results of previous calculations to the values of any next solved optimization problem in the scheme (\ref{eq:9}) without any time-consuming calculations of the values of the criteria $f_i (y)$, $1 \leq j \leq s$, of the problem (\ref{eq:1}) \cite{c30,c31}. 
%new text begin
This possibility of reusing search information can be briefly presented as follows.
%\begin{itemize}
%\item 
%При минимизации очередного критерия $f_l(y(x)), 1\leq l\leq s,$ получаемая поисковая информация накапливается в форме множества $A_k$ из (\ref{eq:21}) (следует отметить, что в $A_k$ наряду со значениями $z_i, 1\leq i\leq k,$ минимизируемого критерия $f_l(x)$ запоминаются и вычисленные значения критериев $f_i, 1\leq i\leq k$.
%\item 
%При переходе к минимизации следующего критерия $f_{l+1}(y(x))$ значения этого критерия $z_i, 1\leq i\leq k,$ во всех точках $x_i, 1\leq i\leq k,$ ранее выполненных итераций глобального поиска могут получены при помощи быстро выполнимых вычислений на основе запомненных значений критериев $f_i, 1\leq i\leq k$ в поисковой информации $A_k$
%\begin{equation}\label{new_equation}
%z'_i=Z(f_i), 1\leq i\leq k,
%\end{equation}
%где процедура вычисления значений $z'_i, 1\leq i\leq k$ определяется правилами (\ref{eq:9}), (\ref{eq:18})--(\ref{eq:20}) без повторных вычислениях значений критериев $f_l(y(x)), 1\leq l\leq s$.
%\end{itemize}
%new text end
This reuse of search information to find the next efficient decision reduces the number of computations needed to solve each next optimization problem to just a few iterations -- see Section \ref{sec:5} for the results of numerical experiments.

Moreover, using the search information $A_k$ allows reducing the multistage solution of the MCOlex problem in accordance with the scheme (\ref{eq:9}) to the solution of the only global optimization problem in the last stage of the scheme (\ref{eq:9}), namely
\begin{equation}\label{eq:30}
P_{\delta lex} (f,\delta,D)=Arg \min_{y \in D} {f_s(y)}, f_j(y)\leq f_j^* + \delta_j, 1 \leq j < s,
\end{equation}
where for the unknown a priori values $f_j^*$, $1 \leq j < s$, the estimates $z_j^k$, ${1 \leq j < s}$, from (\ref{eq:23}) obtained from the search information $A_k$ can be used, and for the concessions $\delta_j$, $1 \leq j < s$, the normalized values can be applied, that is
\begin{equation}\label{eq:31}
	\begin{matrix}
	\delta_j = z_j^k + \theta_j (g_j^k-z_j^k ), 0 \leq \theta_j \leq 1, 1 \leq j < s,\\
	g_j^k = \max{\{ g_j (x_i ) : \nu(x_i)=j, 1 \leq i \leq k \}}.
	\end{matrix}
\end{equation}
%new text begin
%При этом, как и ранее, при любых изменениях оценок  $f_j^*, \delta_j, g_j^k, 1\leq j < s, 1 \leq i \leq k,$ решение задачи (\ref{eq:30}) может быть продолжено с использованием всей имеющейся поисковой информации после быстро выполняемых вычислений новых значений $z'_i,1\leq i \leq k,$ в соответствии с процедурой (\ref{new_equation}).
%new text end
In particular, a similar one-step scheme (\ref{eq:30})--(\ref{eq:31}) can be applied through the use of AGCS.

With a complete scheme for calculating criteria values, i.e. when at the iteration points $x_i$, $1 \leq i \leq k$, the values of all criteria $f_j(x_i)$, $1 \leq j \leq s$, $1 \leq i \leq k$, are calculated, the effect of using the search information $A_k$ can be even more significant. In this case, the search information can also be used when changing the values of the concessions $\delta_j$, $1 \leq j < s$, and thus, the solution of the next problem from the $\mathbb{F}$ family (\ref{eq:10}) can be carried out repeatedly using the results of all previously performed calculations. The results of numerical experiments (see Section \ref{sec:5}), demonstrate that the volume of computations performed, based on reusing search information, can significantly decrease.

The AGCS algorithm, supplemented by the ability to use search information to solve MMCOlex problems, will be referred to below as the global search algorithm for multistage solutions of a set of multiextremal optimization problems with nonlinear constraints (MAGCS).

@@

For MAGCS, the problem (\ref{eq:43}) is considered as the MMCOlex problem, which includes 50 subproblems (\ref{eq:30}) with different values
$\theta_1$ from (\ref{eq:31}), uniformly distributed in the interval $[0,1]$. The accuracy $\varepsilon=0.05$ from (\ref{eq:28}) and the reliability parameters $r_1=r_2=3.0$ from (\ref{eq:29}) were used. The results of the executed experiments are presented in Table \ref{tab:1}.

@@

All the subsequent numerical experiments focused on the solution of 
the MMCOlex problems, which includes 100 two-criteria MCO problems, the efficiency criteria of which were set by the GKLS generator.
Each MCO problem from the generated set is considered as the MMCOlex problem, which includes from 10 to 100 subproblems (\ref{eq:30}) with different
values $\theta_1$ from (\ref{eq:31}), uniformly distributed in the interval $[0,1]$. The obtained results of numerical experiments were averaged by the number of solved MCO problems.

The solution of MCOlex problems was carried out before the stopping criterion of (\ref{eq:28}) was fulfilled, and the MCOlex problems were considered solved correctly if the distance of the solutions to the Pareto set did not exceed the given accuracy $\varepsilon$ from (\ref{eq:28}).

@@

The first column (SP) in the Table \ref{tab:2} contains the total number of MCOlex subproblems to be solved. The PI columns show the average number of iterations for solving one MCO problem. The SPI columns represent the average number of iterations for solve a single MCOlex subproblem. The SPS columns contain a fraction of MCOlex subproblems solved with the required accuracy. The HV and DU columns show the hypervolume (HV) and the distribution uniformity (DU) indices of the calculated approximation of the Pareto set, respectively. The last column (RI) shows the degree of the reduction in the number of global search iterations performed when solving subproblems (\ref{eq:30}) by reusing the search information.

The results presented in Table \ref{tab:2} demonstrate that the use of search information allows to obtain an estimate of the Pareto set after a substantially smaller total number of iterations of the global search (the PI columns). It should also be noted that as the number of solved subproblems increases, the MAGCS allows to build estimates of the Pareto set with better HV and DU indices.

For a better visual representation, Fig. \ref{fig:4} shows the average number of iterations performed to solve a single MCOlex subproblem with and without search information. The presented graphs convincingly demonstrate a significant reduction in the number of global search iterations when solving the MCOlex subproblems using search information.

\begin{figure}
  \centering
  \includegraphics[width=1\linewidth]{fig4}
  \caption{Average number of iterations required to solve one MCOlex subproblem depending on the total number of subproblems solved}
  \label{fig:4}
\end{figure}

Fig. \ref{fig:5} shows the reduction in the number of global search iterations when solving MCOlex subproblems using search information. The graphs presented show that when the number of solved MCOlex subproblems grows linearly, the number of iterations required to solve one subproblem is reduced almost linearly as well.

\begin{figure}
  \centering
  \includegraphics[width=1\linewidth]{fig5}
  \caption{Reduction of the number of iterations performed to solve one MCOlex subproblem depending on the total number of subproblems solved}
  \label{fig:5}
\end{figure}

It should be noted that the number of global search iterations is also reduced in the course of sequential solution of the subproblems of the same MCO problem. Table \ref{tab:3} illustrates this trend showing the average number of iterations required to find a solution to one MCOlex subproblem, with subdividing the sequence of solved subproblems in groups of 10 MCOlex subproblems. The first row (SP) contains the MCOlex subproblem numbers belonging to the individual subproblem groups. The PI rows show the total number of iterations of the global search performed for all previous problems. The rows labeled as ``PI new'' show the number of iterations performed to solve the next group of MCOlex subproblems with the numbers of the problems given in each column heading. The last row labeled as ``FPI new'' shows the proportion of new iterations of global search performed to solve the next group of MCOlex subproblems.

@@

Fig. \ref{fig:6} shows the values of the achieved reduction in the number of additional iterations required to solve the next set of subproblems (the last row of Table \ref{tab:3}). The above data demonstrates that solving the next group of MCOlex subproblems requires about $6\%$ of new global search iterations in addition to the available search information.

\begin{figure}
  \centering
  \includegraphics[width=1\linewidth]{fig6}
  \caption{Proportion of new global search iterations for solving the next group of MCOlex subproblems}
  \label{fig:6}
\end{figure}

The third set of numerical experiments was designed to demonstrate the scalability of the proposed approach by solving more complex two-criteria four-dimensional MCO problems under the same conditions as in the previous series of calculations. The results of the experiments are presented in Tables \ref{tab:4}--\ref{tab:5} and in Fig. \ref{fig:4}--\ref{fig:6}. These data demonstrate that the experiments carried out to solve more complex MCO problems confirm the conclusions made earlier -- for example, solving the next group of MCOlex subproblems requires, in addition to the available search information, from $0.5\%$ to $14.2\%$ more global search iterations (Table \ref{tab:5}, last row).

@@

In this paper a new approach is proposed for solving computationally expensive lexicographic multicriteria optimization problems (MCOlex), which can be complicated by the multiextremal nature of the efficiency criteria and extensive volume of computations required to calculate the criteria values. A key feature of this class of problems is the ability to alter the ranking of efficiency criteria in terms of importance during calculations, which in turn necessitates a multistage solution for MCOlex problems.  This new type of problems is referred to as MMCOlex problems.

The computational complexity of addressing the formulated MMCOlex problems can be overcome by solving a sequence of global optimization problems with nonlinear constraints using efficient information-statistical methods of global optimization. A core element in this approach is the ability to use all the search information obtained in the course of solving the MMCOlex problems. With each new stage of the solution, this search information allows incorporating the previously calculated values of efficiency criteria into the values of the next scalar problem of multiextremal optimization. The search information retrieved becomes part of the current optimization state and, through optimization methods, results in adaptive planning of the iterations to run the global search.

The results of numerical experiments demonstrate that this approach can sufficiently reduce the computational complexity of MMCOlex problem-solving.
